---
title: "Effectiveness of Ferritin‐guided Donation Intervals in Blood Donors: results of the Stepped Wedge Cluster‐randomized FIND’EM Trial"
format:
  html:
    code-fold: true
#    page-layout: full
    toc: true
    embed-resources: true
---

```{r, collapse=TRUE}
#| label: load-packages

library(tidyverse)       
library(haven)           # importing SPSS data files
library(furniture)       # nice tables of descriptives
library(corrplot)        # visualize correlations
library(ggplot2)         # plot making
library(GGally)          # extensions to ggplot2
library(sjlabelled)      # work with SPSS data
library(lubridate)       # dates
library(lme4)            # mixed models
library(table1)          # for producing t1
library(sjPlot)          # for model outputs
library(car)             # to calculate VIF
library(broom)           # for function augment
library(stringr)
library(reshape)
library(cli)
library(forestmangr)
library(cowplot)
library(ggpubr)
library(robustbase)      # robust regression
library(data.table)
library(dplyr)
library(broom.mixed)
library(sjstats)



setwd("~/Amber/FINDEM") 
```

# Data

We load the data and present a summary.

```{r, collapse = T}
#| label: read data

data_raw <- readRDS("~/Amber/Data/FINDEM/FINDEM_data.rds")
#clean data
data_clean <- data_raw%>% 
  mutate(HbDef = factor(case_when(HbGoedgekeurd == 0 ~ 1, HbGoedgekeurd == 1 ~ 0))) %>% 
  mutate(Gender = factor(Geslacht) %>% fct_recode("Male" = "Man", "Female" = "Vrouw")) %>% 
  mutate(Timepoint = case_when(Meetweek == 0 ~ as.Date('2017-11-01'), Meetweek == 1 ~ as.Date('2019-03-01'), Meetweek == 2 ~ as.Date('2019-11-01'), TRUE ~ as.Date('2020-11-23'))) %>% 
  mutate(Step = case_when(stap2 == 1.0 ~ as.Date('2017-11-01'), stap2 == 2.1 ~ as.Date('2018-05-01'), stap2 == 2.2~ as.Date('2018-10-01'), stap2 == 3.1 ~ as.Date('2019-03-01'), stap2 == 3.2 ~ as.Date('2019-10-01'), stap2 == 4.0 ~ as.Date('2019-11-01'))) %>% 
  mutate(Intervention_months = interval(Step, Timepoint) %/% months(1)) %>% 
  mutate(CalendarTime = interval(as.Date('2017-11-01'), Timepoint)%/% months(1))%>%
  mutate(Intervention_months =  if_else(Intervention_months < 0, 0, Intervention_months)) %>% 
  mutate(Intervention = factor(case_when(Intervention_months == 0 ~ 0, TRUE & !is.na(Intervention_months)~ 1)))%>% 
  mutate(PostMeno = factor(case_when(Gender == "Female" & LeeftijdBijDOnatie < 45 ~ 0, Gender== "Female" & LeeftijdBijDOnatie >= 45 ~ 1))) %>%
  dplyr::rename(Ferritin = Ferritine_FINDEM, Age = LeeftijdBijDOnatie, Weight = Gewicht, Height = Lengte)  %>% 
  mutate(IronDef = factor(case_when(Ferritin >= 15 ~ 0, Ferritin < 15 ~ 1)))%>% 
  dplyr::select(-HbGoedgekeurd, -Geslacht)%>%
  mutate(Intervention_months2 =  if_else(Intervention_months < 6, 0, Intervention_months))%>%
  mutate(Intervention_months_factor = factor(case_when(Intervention_months == 0 ~ 0, Intervention_months == 1 ~ 0, Intervention_months == 5 ~ 0, Intervention_months == 8 ~ 1, Intervention_months == 10 ~ 1, Intervention_months == 12 ~ 2, Intervention_months == 13 ~ 2,Intervention_months == 16 ~ 2,Intervention_months == 18 ~ 3, Intervention_months == 20 ~ 3, Intervention_months == 24 ~ 4, Intervention_months == 25 ~ 4, Intervention_months == 30 ~ 5, Intervention_months == 36 ~ 6))) %>%
    mutate(Intervention_months_factor2 = factor(case_when(Intervention_months == 0 ~ 0, Intervention_months == 1 ~ 1, Intervention_months == 5 ~ 1, Intervention_months == 8 ~ 2, Intervention_months == 10 ~ 2, Intervention_months == 12 ~ 3, Intervention_months == 13 ~ 3,Intervention_months == 16 ~ 4,Intervention_months == 18 ~ 4, Intervention_months == 20 ~ 4, Intervention_months == 24 ~ 5, Intervention_months == 25 ~ 5, Intervention_months == 30 ~ 6, Intervention_months == 36 ~ 7))) %>%
  mutate(COVID = factor(ifelse(Donatiedatum >= "2019-10-01" & Donatiedatum <= "2020-02-01", 1,0))) %>% 
  mutate(Hb = Hb/0.6206)%>%
  mutate(Hb = replace(Hb, Hb == (999/0.6206), NA))


## Quick view
summary(data_clean)


data_scale <- data_clean
data_scale$Age <- scale(data_scale$Age, center=TRUE, scale = FALSE)
data_scale$Weight <- scale(data_scale$Weight, center=TRUE, scale = FALSE)
data_scale$Height <- scale(data_scale$Height, center=TRUE, scale = FALSE)
data_scale$FerritinLog10 <- log10(data_scale$Ferritin) 

```

# Table 1

See section Descriptives of main FINDEM.html file for the study population description.
Description of the general donor population:

```{r}
#we select the general donor population as all whole blood and new donors, 1 year before the start of the study
#variables we have to make are: total number of donations, donations in the past 2 years, return(?)
pss2date <- function(x) as.Date(x/86400, origin = "1582-10-14") 

donaties <- readRDS("~/Amber/Data/Algemeen/alledonaties_2008_2022.rds") 
donaties <- donaties %>% mutate(Donatiedatum = pss2date(Donatiedatum), record = 1) %>% filter(Donatiesoortcode == "Volbloed"|Donatiesoortcode=="Nieuwe donorkeuring")%>% arrange(Donatiedatum) %>% group_by(KeyID) %>% mutate(TotaalOpBaseline=sum(record)) %>% select(-record)

#test from other script
donaties <- donaties[order(as.Date(donaties$Donatiedatum, format="%Y-%m-%d")),]
donaties$donatienr <- with(donaties, ave(rep(1, nrow(donaties)), KeyID, FUN = seq_along))

donaties <- transform(donaties, id = as.numeric(factor(KeyID)))

donaties<-setDT(donaties)

donaties[, n:= 1:.N - findInterval(Donatiedatum - (2*365), Donatiedatum), by=.(KeyID)]
donaties <- donaties %>% dplyr::rename(count24 = n)

donaties <- as.data.frame(donaties)
#

donaties_gp <- donaties %>% filter(Donatiedatum > "2016-09-11") %>% arrange(KeyID, Donatiedatum)

#create pointer for next donation
idx<-1:nrow(donaties_gp)
precursor<-idx+1
precursor[nrow(donaties_gp)]<-1

# calculate time since last donation
donaties_gp$dt<-NA
donaties_gp$dt<-as.numeric(donaties_gp$Donatiedatum[precursor]-donaties_gp$Donatiedatum)
donaties_gp$dt[donaties_gp$KeyID != donaties_gp$KeyID[precursor]]<-NA
donaties_gp$dt[is.na(donaties_gp$dt)] <- as.numeric(Sys.Date()-donaties_gp$Donatiedatum)

#return on time?
donaties_gp$return <- NA
donaties_gp$return[donaties_gp$Geslacht=="M" & donaties_gp$dt<(180+56)] <- 1
donaties_gp$return[donaties_gp$Geslacht=="F" & donaties_gp$dt<(180+122)] <- 1
donaties_gp$return[donaties_gp$Ferritine<15 & donaties_gp$dt<(180+365)] <- 1
donaties_gp$return[donaties_gp$Ferritine<=15 & donaties_gp$Ferritine<30 & donaties_gp$dt<(180+180)] <- 1
donaties_gp$return[donaties_gp$HbGoedgekeurd=="afgekeurd" & donaties_gp$dt<(90+180)] <- 1
donaties_gp$return[is.na(donaties_gp$return)]<-0

donaties_gp <- donaties_gp %>% filter(Donatiedatum < "2017-09-11")
donaties_gp <- donaties_gp %>% mutate(Geboortedatum = pss2date(Geboortedatum), Age = as.numeric(Donatiedatum - Geboortedatum)/365.25)

donors <- readRDS(file = "~/Amber/Data/Algemeen/DonorsTotaal.rds") %>% select(KeyID, Lengte, Gewicht)

genpop <- merge(donaties_gp, donors, by = "KeyID")

genpop <- genpop %>% mutate(HbDef = factor(case_when(HbGoedgekeurd == "afgekeurd" ~ 1, HbGoedgekeurd == "goedgekeurd" ~ 0)), Ndonation = factor(case_when(Donatiesoortcode == "Nieuwe donorkeuring" ~ 1, Donatiesoortcode == "Volbloed" ~ 0)), Hb = Hb/0.6206) %>% dplyr::rename(Ferritin = Ferritine, Height = Lengte, Weight = Gewicht, Previous2yr = count24)
```

```{r}
prefemalest1<-table1::table1(~ Age + Hb + Ferritin + Height + Weight + TotaalOpBaseline + Previous2yr + factor(HbDef) + factor(return) + factor(Ndonation), data = subset(genpop, genpop$Geslacht=="F" & genpop$Age< 45), caption = "Premenopausal females", render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"), render.categorical=c(.="FREQ (PCT)"))
prefemalest1
```
```{r}
postfemalest1<-table1::table1(~ Age + Hb + Ferritin + Height + Weight + TotaalOpBaseline + Previous2yr + factor(HbDef) + factor(return) + factor(Ndonation), data = subset(genpop, genpop$Geslacht=="F" & genpop$Age>=45), caption = "Postmenopausal females", render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"), render.categorical=c(.="FREQ (PCT)"))
postfemalest1
```
```{r}
malest1<-table1::table1(~ Age + Hb + Ferritin + Height + Weight + TotaalOpBaseline + Previous2yr + factor(HbDef) + factor(return) + factor(Ndonation), data = subset(genpop, genpop$Geslacht=="M"), caption = "Males", render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"), render.categorical=c(.="FREQ (PCT)"))
malest1
```

# Table 2

See section Analysis \> Linear regression of main FINDEM.html file and section Analysis of FINDEM_questionnaire.html file.


# Figure 1

Figure 1 was produced in Microsoft Visio.

# Figure 2

Figure 2 was produced in Microsoft Visio.

# Figure 3

```{r, warnings=F, errors=F, messages=F}
data_clean$LowFerr <- 0
data_clean$LowFerr[data_clean$Ferritin>=15 & data_clean$Ferritin < 30] <- 1

figure2M <- data.frame(matrix(NA,    # Create empty data frame
                          nrow = 7, # we make a dataset per sex and then have one row per month group of the intervention
                          ncol = 4)) #we need columns for the percentage hb deferral, iron deficiency and low ferritin
colnames(figure2M) <- c("Months", "Hb", "ID", "LowFer")
figure2M$Months <- c("Control", "6-11", "12-17", "18-23", "24-29", "30-35", "36+")

figure2PreF <- data.frame(matrix(NA,    # Create empty data frame
                          nrow = 7, # we make a dataset per sex and then have one row per month group of the intervention
                          ncol = 4)) #we need columns for the percentage hb deferral, iron deficiency and low ferritin
colnames(figure2PreF) <- c("Months", "Hb", "ID", "LowFer")
figure2PreF$Months <- c("Control", "6-11", "12-17", "18-23", "24-29", "30-35", "36+")

figure2PostF <- data.frame(matrix(NA,    # Create empty data frame
                          nrow = 7, # we make a dataset per sex and then have one row per month group of the intervention
                          ncol = 4)) #we need columns for the percentage hb deferral, iron deficiency and low ferritin
colnames(figure2PostF) <- c("Months", "Hb", "ID", "LowFer")
figure2PostF$Months <- c("Control", "6-11", "12-17", "18-23", "24-29", "30-35", "36+")

for (i in 1:7){
  figure2M[i,2] <- (length(which(data_clean$Gender == "Male" & data_clean$Intervention_months_factor == (i-1) & data_clean$HbDef==1))/length(which(data_clean$Gender == "Male" & data_clean$Intervention_months_factor == (i-1))))*100
  figure2M[i,3] <- (length(which(data_clean$Gender == "Male" & data_clean$Intervention_months_factor == (i-1) & data_clean$IronDef==1))/length(which(data_clean$Gender == "Male" & data_clean$Intervention_months_factor == (i-1))))*100
    figure2M[i,4] <- (length(which(data_clean$Gender == "Male" & data_clean$Intervention_months_factor == (i-1) & data_clean$LowFerr ==1))/length(which(data_clean$Gender == "Male" & data_clean$Intervention_months_factor == (i-1))))*100
    
    mdat = melt(figure2M, id.vars=c("Months"),
            measure.vars=c("ID", "LowFer"))
    mdat$Months <- factor(mdat$Months, levels = c("Control", "6-11", "12-17", "18-23", "24-29", "30-35", "36+"))
    mdat <- merge(mdat, figure2M[,1:2], by = "Months")
    mdat[c(2,4,6, 8, 10,14), 4]<-NA
    mdat$Sex <- "Males"
    
      figure2PreF[i,2] <- (length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 0 & data_clean$Intervention_months_factor == (i-1) & data_clean$HbDef==1))/length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 0 & data_clean$Intervention_months_factor == (i-1))))*100
  figure2PreF[i,3] <- (length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 0 & data_clean$Intervention_months_factor == (i-1) & data_clean$IronDef==1))/length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 0 & data_clean$Intervention_months_factor == (i-1))))*100
    figure2PreF[i,4] <- (length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 0 & data_clean$Intervention_months_factor == (i-1) & data_clean$LowFerr ==1))/length(which(data_clean$Gender == "Female" &  data_clean$PostMeno == 0 & data_clean$Intervention_months_factor == (i-1))))*100
    
    mdatPreF = melt(figure2PreF, id.vars=c("Months"),
            measure.vars=c("ID", "LowFer"))
    mdatPreF$Months <- factor(mdatPreF$Months, levels = c("Control", "6-11", "12-17", "18-23", "24-29", "30-35", "36+"))
    mdatPreF <- merge(mdatPreF, figure2PreF[,1:2], by = "Months")
    mdatPreF[c(2,4,6, 8, 10,14), 4]<-NA
    mdatPreF$Sex <- "Premenopausal females"
          figure2PostF[i,2] <- (length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 1 & data_clean$Intervention_months_factor == (i-1) & data_clean$HbDef==1))/length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 1 & data_clean$Intervention_months_factor == (i-1))))*100
  figure2PostF[i,3] <- (length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 1 & data_clean$Intervention_months_factor == (i-1) & data_clean$IronDef==1))/length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 1 & data_clean$Intervention_months_factor == (i-1))))*100
    figure2PostF[i,4] <- (length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 1 & data_clean$Intervention_months_factor == (i-1) & data_clean$LowFerr ==1))/length(which(data_clean$Gender == "Female" &  data_clean$PostMeno == 1 & data_clean$Intervention_months_factor == (i-1))))*100
    
    mdatPostF = melt(figure2PostF, id.vars=c("Months"),
            measure.vars=c("ID", "LowFer"))
    mdatPostF$Months <- factor(mdatPostF$Months, levels = c("Control", "6-11", "12-17", "18-23", "24-29", "30-35", "36+"))
    mdatPostF <- merge(mdatPostF, figure2PostF[,1:2], by = "Months")
    mdatPostF[c(2,4,6, 8, 10,14), 4]<-NA
    mdatPostF$Sex <- "Postmenopausal females"

}

mdat_total <- rbind(mdat, mdatPreF, mdatPostF)
```

```{r, warning=FALSE, error=FALSE, message=FALSE}
#| label: make figures

# Function to scale secondary axis
scale_function <- function(x, scale, shift){
  return ((x)*scale - shift)
}

# Function to scale secondary variable values
inv_scale_function <- function(x, scale, shift){
  return ((x + shift)/scale)
}
sex <- c("Males", "Premenopausal females", "Postmenopausal females")

for(i in sex){
  max_first  <- max(mdat_total$value)+20   # Specify max of first y axis
max_second <- max(mdat_total$Hb, na.rm=T) # Specify max of second y axis
min_first  <- min(mdat_total$value)+20   # Specify min of first y axis
min_second <- min(mdat_total$Hb, na.rm=T) # Specify min of second y axis

# scale and shift variables calculated based on desired mins and maxes
scale = max_second/max_first
shift = 0 #min_first - min_second

plot <- ggplot(data=mdat_total[mdat_total$Sex==i,]) + 
  geom_bar(aes(x=Months, y=value, fill = variable), stat="identity", width = 0.4, position = "stack") + 
  xlab("Time since implementation (months)")+ 
  ylab("Percentage")+
  geom_bar(aes(x=Months, y=inv_scale_function(Hb, scale, shift)), stat = "identity", position = position_nudge(x = 0.5), width = 0.4, col = "grey", fill = "grey")+
  ggtitle(paste0(i)) + 
  scale_y_continuous(name = "Donors with low ferritin (%)", sec.axis = sec_axis(~scale_function(., scale, shift), name="Donors deferred based on Hb (%)"))+ 
  theme(axis.title.y = element_text(color = "black", size=10),
        axis.title.y.right = element_text(color = "black", size=10))
print(plot)
}
```

```{r, warning=F, message=F}

order <- c("Premenopausal females", "Postmenopausal females", "Males")
mdat_total <- mdat_total %>% mutate(Sex = ordered(Sex, levels = order))
plot <- ggplot(data=mdat_total) + 
  geom_bar(aes(x=Months, y=value, fill = variable), stat="identity", width = 0.4, position = "stack") + scale_fill_manual(values = c("ID" = "#ff9700", "LowFer"= "#1cdbfc"), labels = c("Deficient (<15 ng/mL)", "Low (<30 ng/mL)"), name="Ferritin")+
  geom_text(aes(x=Months, y=value, label = round(value,1), fill = variable), 
            position = position_stack(vjust = .5), size=2)+
  xlab("Time since implementation (months)")+ 
  ylab("Percentage")+
  geom_bar(aes(x=Months, y=inv_scale_function(Hb, scale, shift), color = "grey"), stat = "identity", position = position_nudge(x = 0.5), width = 0.35, fill = "grey") + scale_color_manual(values = c("grey" = "grey"), labels = c("Hb < 12.5/13.5 g/dL (F/M)"), name="Deferral")+
  geom_text(aes(x=Months, y=inv_scale_function(Hb, scale, shift)/2, label = round(Hb,1), fill = variable), 
            position = position_nudge(x = 0.5), size=2)+
  ggtitle("Donors presenting with low ferritin and hemoglobin")+
  scale_y_continuous(name = "Donors with low ferritin (%)", sec.axis = sec_axis(~scale_function(., scale, shift), name="Donors deferred based on Hb (%)"))+ 
  theme(axis.title.y = element_text(color = "black", size=10),
        axis.title.y.right = element_text(color = "black", size=10))+
  facet_grid(.~ Sex) + 
  theme(panel.background = element_rect(fill = "white",
                                    colour = "white",
                                    size = 0.5, linetype = "solid"),
    panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                    colour = "#ededed"), 
    panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                    colour = "#ededed"))

ggsave(plot, width = 12, height = 5, file = "~/Amber/FINDEM/figure2.jpg")
ggsave(plot, width = 12, height = 5, file = "~/Amber/FINDEM/figure2.pdf")
ggsave(filename = "figure2.eps",width = 12, height = 5,
       plot = print(plot),
       device = "eps")
```

# Forest plot primary outcomes

Load all models:
```{r}
#| label: forest plot

hb_fit_M <- lm(Hb ~ Age + Weight + Height + CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male")
hb_fit_preF <- lm(Hb ~ Age + Weight + Height +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0)
hb_fit_postF <- lm(Hb ~ Age + Weight + Height + CalendarTime +Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1)

ferritin_fit_M <- lm(FerritinLog10 ~ Age + Weight + Height +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male")
ferritin_fit_preF <- lm(FerritinLog10 ~ Age + Weight + Height +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0)
ferritin_fit_postF <- lm(FerritinLog10 ~ Age + Weight + Height +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1)

hbdef_fit_M <- glm(formula = HbDef ~ Age + Weight + Height +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male" ,family = binomial)
hbdef_fit_preF <- glm(formula = HbDef ~ Age + Weight + Height +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0,family = binomial)
hbdef_fit_postF <- glm(formula = HbDef ~ Age + Weight + Height +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1,family = binomial)

ID_fit_M <- glm(formula = IronDef ~ Age + Weight + Height +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male" ,family = binomial)
ID_fit_preF <- glm(formula = IronDef ~ Age + Weight + Height +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0,family = binomial)
ID_fit_postF <- glm(formula = IronDef ~ Age + Weight + Height +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1,family = binomial)

models <- ls()
models <- models[grepl("fit", models)]

load_models <- function(model){
  info <- str_split(model, "_", simplify=T)
  if(grepl("hbdef", model)){coef <- tidy(get(model), conf.int=T, exponentiate=T)
  } else if  (grepl("ID", model)){coef <- tidy(get(model), conf.int=T, exponentiate=T)
  } else if (grepl("return", model)){coef <- tidy(get(model), conf.int=T, exponentiate=T)} else {coef <- tidy(get(model), conf.int=T)}
  coef <- coef %>% mutate(outcome = info[1], sex = info[3]) %>% filter(grepl("Intervention", term))
}

order <- rev(c("Intervention_months_factor1", "Intervention_months_factor2", "Intervention_months_factor3", "Intervention_months_factor4", "Intervention_months_factor5", "Intervention_months_factor6"))

datalist <- lapply(models, load_models)
data.forestplot <- bind_rows(datalist) %>% mutate(term = ordered(term, levels = order)) %>% mutate(sex = dplyr::recode(sex, "M" = "Males", "postF" = "Postmenopausal\nfemales", "preF" = "Premenopausal\nfemales"), sex = ordered(sex, levels = c("Premenopausal\nfemales", "Postmenopausal\nfemales", "Males")), outcome = dplyr::recode(outcome, "ferritin" = "Ferritin", "hb" = "Hemoglobin", "hbdef" = "Hb based \ndeferral", "ID" = "Iron \ndeficiency"),  significant = as.factor(ifelse(p.value <= (0.05/4), 1, 0)), significant = ordered(significant, levels = c("1", "0")))

p1 <- ggplot(data.forestplot[data.forestplot$outcome=="Ferritin" | data.forestplot$outcome == "Hemoglobin",], aes(x=term, y=estimate, shape = significant)) +
         geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
                       width = 0.2,size  = 0.5,
                       position = "dodge", color="#1cdbfc") +
  geom_hline(yintercept = 0, color = "#ff9700", linewidth = 0.35) +
  scale_shape_manual(values = c(16, 1)) + 
  ylim(-0.2,0.5) +
  geom_point() + coord_flip() + scale_x_discrete(breaks = c("Intervention_months_factor1", "Intervention_months_factor2", "Intervention_months_factor3", "Intervention_months_factor4", "Intervention_months_factor5", "Intervention_months_factor6"),
                           labels = c("6-11", "12-17", "18-23", "24-29", "30-35", "36+")) + ylab("Regression coefficient") + facet_grid(outcome~sex) + theme(legend.position = "none",axis.title.y=element_blank())

p2 <- ggplot(data.forestplot[data.forestplot$outcome=="Iron \ndeficiency" | data.forestplot$outcome == "Hb based \ndeferral" | data.forestplot$outcome == "Return",], aes(x=term, y=estimate, shape = significant)) +
         geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
                       width = 0.2,size  = 0.5,
                       position = "dodge", color="#1cdbfc") +
  geom_hline(yintercept = 1, color = "#ff9700", linewidth = 0.35) +
  scale_shape_manual(values = c(16, 1), name = "Statistical \nsignificance \n(p<0.0125)        ", labels = c("Yes", "No")) +
  geom_point() + coord_flip() + scale_x_discrete(breaks = c("Intervention_months_factor1", "Intervention_months_factor2", "Intervention_months_factor3", "Intervention_months_factor4", "Intervention_months_factor5", "Intervention_months_factor6"),
                           labels = c("6-11", "12-17", "18-23", "24-29", "30-35", "36+")) + ylab("Odds ratio")  + facet_grid(outcome~sex) + theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.key.height = unit(0.5, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), legend.position = "none", axis.title.y = element_blank(), legend.title=element_text(size=7), legend.text=element_text(size=7)) #change legend key width



figure <- plot_grid(p1, p2, ncol = 2, align = 'h', rel_widths = c(15/30, 15/30))
legend <- get_legend(
  # create some space to the left of the legend
  p2 + theme(legend.position = "right", legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.key.height = unit(0.5, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), axis.title.y = element_blank(), legend.title=element_text(size=7), legend.text=element_text(size=7))
)
figure <- plot_grid(figure, legend, align = "h", ncol = 2, rel_widths = c(2, 0.2))
annotate_figure <- annotate_figure(figure,
                top = text_grob("Associations of implementation of ferritin-guided donation \nintervals with primary study outcomes", color = "black", size = 14),
                bottom = text_grob("Ferritin: log10 ng/mL; Hb: g/dL; Iron deficiency: ferritin < 15 ng/mL; Hb deferral: males < 13.5 g/dL, females < 12.5 g/dL. \nReference group: no implementation and 0-5 months since implementation. \nAll analyses are adjusted for age, height, weight and time. \nStatistical significance: Bonferroni correction applied to adjust for multiple testing (0.05/4).", color = "black",
                                   hjust = 0, x = 0, size = 5),
                left = text_grob("Time after intervention (months)", color = "black", rot = 90, vjust=1, x=0.1)
                )
ggsave(annotate_figure, file = "forestplot.jpg", width = 10, height = 4)
```

# Forest plot secondary outcomes
Load data:
```{r}
#| label: load questionnaire data
returnOT_fitQ_M <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + COVID +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male",family = binomial)
returnOT_fitQ_preF <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + COVID +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0,family = binomial)
returnOT_fitQ_postF <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + COVID +CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1,family = binomial)


data_clean <- read_sav("~/Amber/Data/FINDEM/Final_survey.sav") 
data_clean <- data_clean %>% dplyr::select(-Einnummer, -Cluster, -Date, -Meetweek) %>% 
  mutate(Gender = factor(Gender) %>% fct_recode("Male" = "1", "Female" = "2")) %>%
  mutate(PICA = factor(PICA) %>% fct_recode("Yes" = "1", "No" = "0")) %>% 
  mutate(RLS = factor(case_when(RLS == 2 ~ 1, RLS == 1 ~ 0)) %>% fct_recode("Yes" = "1", "No" = "0")) %>% 
  mutate(Intervention_months_factor = factor(case_when(Intervention_months_s < 6 ~ 0, Intervention_months_s >= 6 & Intervention_months_s <12 ~ 1, Intervention_months_s >= 12 & Intervention_months_s < 18 ~ 2, Intervention_months_s >= 18  & Intervention_months_s < 24 ~ 3,Intervention_months_s >= 24 ~ 4))) %>% 
  mutate(CalendarTime = interval(as.Date('2017-11-01'), Timepoint)%/% months(1))%>%
  mutate(Fergroup = factor(case_when(Ferritin < 15 ~ 1, Ferritin >= 15 & Ferritin <= 30~ 2, Ferritin > 30 ~ 3)) %>%  fct_recode("Ferritin < 15" = "1", "Ferritin 15-30" = "2", "Ferritin > 30" = "3")) %>% 
  mutate(Warmglow = as.numeric(Warmglow))  %>% mutate(Warmglow = as.factor(case_when(Warmglow >=6 ~ 1, TRUE & !is.na(Warmglow) ~ 0)) %>% fct_recode("Yes" = "1", "No" = "0"))

data_scale <- data_clean
data_scale$Age <- scale(data_scale$Age, center=TRUE, scale = FALSE)
data_scale$Weight <- scale(data_scale$Weight, center=TRUE, scale = FALSE)
data_scale$Height <- scale(data_scale$Height, center=TRUE, scale = FALSE)
```

Load models:
```{r}
PICA_fitQ_M <- glm(formula = PICA ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male" ,family = binomial)
PICA_fitQ_preF <- glm(formula = PICA ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0 ,family = binomial)
PICA_fitQ_postF <- glm(formula = PICA ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1 ,family = binomial)

RLS_fitQ_M <- glm(formula = RLS ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male" ,family = binomial)
RLS_fitQ_preF <- glm(formula = RLS ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0 ,family = binomial)
RLS_fitQ_postF <- glm(formula = RLS ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1 ,family = binomial)

CFQ_fitQ_preF<- lmrob(CFQ_Total ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0)
CFQ_fitQ_postF <- lmrob(CFQ_Total ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1)
CFQ_fitQ_M <- lmrob(CFQ_Total ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male")

SFph_fitQ_preF <- lmrob(SF_ph ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale, subset  = data_scale$Gender == "Female" & data_scale$PostMeno == 0)
SFph_fitQ_postF <- lmrob(SF_ph ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale,  subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1)
SFph_fitQ_M <- lmrob(SF_ph ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male")

SFmh_fitQ_preF <- lmrob(SF_mh ~ Age + Weight + Height+CalendarTime + Intervention_months_factor , data = data_scale, subset  = data_scale$Gender == "Female" & data_scale$PostMeno == 0)
SFmh_fitQ_postF <- lmrob(SF_mh ~ Age + Weight + Height+CalendarTime + Intervention_months_factor , data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1)
SFmh_fitQ_M <- lmrob(SF_mh ~ Age + Weight + Height+CalendarTime + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male")

WG_fitQ_preF <- glm(Warmglow ~ Age + Weight + Height+CalendarTime + Intervention_months_factor , data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0,family = binomial)
WG_fitQ_postF <- glm(Warmglow ~ Age + Weight + Height+CalendarTime + Intervention_months_factor , data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1,family = binomial)
WG_fitQ_M <- glm(Warmglow ~ Age + Weight + Height+CalendarTime + Intervention_months_factor , data = data_scale, subset = data_scale$Gender == "Male" ,family = binomial)

CIS_fitQ_preF <- lmrob(CIS_Totalmean ~ Age + Weight + Height+CalendarTime + Intervention_months_factor , data = data_scale, subset  = data_scale$Gender == "Female" & data_scale$PostMeno == 0)
CIS_fitQ_postF <- lmrob(CIS_Totalmean ~ Age + Weight + Height+CalendarTime + Intervention_months_factor , data = data_scale, subset  = data_scale$Gender == "Female" & data_scale$PostMeno == 1)
CIS_fitQ_M <- lmrob(CIS_Totalmean ~ Age + Weight + Height+CalendarTime + Intervention_months_factor , data = data_scale, subset  = data_scale$Gender == "Male")

modelsQ <- ls()
modelsQ <- modelsQ[grepl("fitQ", modelsQ)]

load_models_Q <- function(model){
  info <- str_split(model, "_", simplify=T)
  if(grepl("WG", model)){coef <- tidy(get(model), conf.int=T, exponentiate=T)
  } else if  (grepl("RLS", model)){coef <- tidy(get(model), conf.int=T, exponentiate=T)
  } else if (grepl("PICA", model)){coef <- tidy(get(model), conf.int=T, exponentiate=T)
  } else if (grepl("return", model)){coef <- tidy(get(model), conf.int=T, exponentiate=T) 
  } else {coef <- tidy(get(model), conf.int=T)}
  coef <- coef %>% mutate(outcome = info[1], sex = info[3]) %>% filter(grepl("Intervention", term))
}

order <- rev(c("Intervention_months_factor1", "Intervention_months_factor2", "Intervention_months_factor3", "Intervention_months_factor4", "Intervention_months_factor5", "Intervention_months_factor6"))

datalist <- lapply(modelsQ, load_models_Q)
data.forestplot <- bind_rows(datalist) %>% mutate(term = ordered(term, levels = order)) %>% mutate(sex = dplyr::recode(sex, "M" = "Male", "postF" = "Postmenopausal\nfemales", "preF" = "Premenopausal\nfemales"), outcome = dplyr::recode(outcome, "WG" = "Warm\nglow", "PICA" = "Pica", "RLS" = "Restless \nlegs syn-\ndrome", "CFQ" = "Cognitive \nfuncti- \noning", "CIS" = "Fatigue", "SFph" = "Physical \nhealth", "SFmh"="Mental\nhealth", "returnOT" = "Return"), outcome = ordered(outcome, levels = c("Return", "Warm\nglow", "Pica", "Restless \nlegs syn-\ndrome","Cognitive \nfuncti- \noning", "Fatigue", "Physical \nhealth", "Mental\nhealth")), significant = as.factor(ifelse(p.value <= 0.05, 1, 0)), significant = ordered(significant, levels = c("1", "0"))) %>% mutate(low_l = ifelse(conf.low < -2, -1.8, NA), high_l = ifelse(conf.high > 2, 1.8, NA))

p1Q <- ggplot(data.forestplot[data.forestplot$outcome=="Cognitive \nfuncti- \noning" | data.forestplot$outcome == "Fatigue" | data.forestplot$outcome == "Physical \nhealth" | data.forestplot$outcome=="Mental\nhealth",], aes(x=term, y=estimate, shape = significant)) +
         geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
                       width = 0.2,size  = 0.5,
                       position = "dodge", color="#1cdbfc") +
  geom_hline(yintercept = 0, color = "#ff9700", linewidth = 0.35) +
  scale_shape_manual(values = c(1, 16)) +
  geom_point() + coord_flip() + scale_x_discrete(breaks = c("Intervention_months_factor1", "Intervention_months_factor2", "Intervention_months_factor3", "Intervention_months_factor4"),
                           labels = c("6-11", "12-17", "18-23", "24-29")) + ylab("Regression coefficient") + facet_grid(outcome~sex) + theme(legend.position = "none",axis.title.y=element_blank(), axis.text.x = element_text(size=6))

p2Q <- ggplot(data.forestplot[data.forestplot$outcome=="Warm\nglow" | data.forestplot$outcome == "Restless \nlegs syn-\ndrome" | data.forestplot$outcome == "Pica" | data.forestplot$outcome=="Return",], aes(x=term, y=estimate, shape = significant)) +
         geom_errorbar(aes(ymin=conf.low, ymax=conf.high), 
                       width = 0.2,size  = 0.5,
                       position = "dodge", color="#1cdbfc") +
  geom_hline(yintercept = 1, color = "#ff9700", linewidth = 0.35) +
  scale_shape_manual(values = c(16, 1), name = "Statistical \nsignificance \n(p<0.05)        ", labels = c("Yes", "No")) +
  geom_point() + coord_flip(ylim = c(0,2)) + scale_x_discrete(breaks = c("Intervention_months_factor1", "Intervention_months_factor2", "Intervention_months_factor3", "Intervention_months_factor4", "Intervention_months_factor5", "Intervention_months_factor6"),
                           labels = c("6-11", "12-17", "18-23", "24-29", "30-35", "36+")) + ylab("Odds ratio")  + facet_grid(outcome~sex) + 
  theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.key.height = unit(0.5, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), legend.position = "none", axis.title.y = element_blank(), legend.title=element_text(size=7), legend.text=element_text(size=7), axis.text.x = element_text(size=6))+ #change legend key width
geom_segment(aes(x = term, xend = term, y = high_l+0.1, yend = high_l+0.3), arrow = arrow(length = unit(0.1, "npc")), col = "#1cdbfc")


figure <- plot_grid(p1Q, p2Q, align = "h", ncol = 2, rel_widths = c(15/30, 15/30))
legend <- get_legend(
  # create some space to the left of the legend
  p2Q + theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.key.height = unit(0.5, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), axis.title.y = element_blank(), legend.title=element_text(size=7), legend.text=element_text(size=7), legend.position = "right")
)
figure <- plot_grid(figure, legend, align = "h", ncol = 2, rel_widths = c(2, 0.2))
annotate_figureQ <- annotate_figure(figure,
                top = text_grob("Associations of implementation of ferritin-guided donation \nintervals with secondary study outcomes", color = "black", size = 14),
                bottom = text_grob("Return: within 6 months of first allowed date. \nCognitive functioning: higher score indicates worse result. \nPhysical & mental health: higher score indicates better result. \nAll analyses are adjusted for age, height, weight and time. \nReturn is additionally adjusted for whether return would be during first COVID lockdown (March-May 2020).", color = "black",
                                   hjust = 0, x = 0, size = 5),
                left = text_grob("Time after intervention (months)", color = "black", rot = 90, vjust=1, x=0.1)
                )
ggsave(annotate_figureQ, file = "forestplot_Q.jpg", width = 10, height = 5)
```

# Combined forest plot

```{r}
final_figure <- plot_grid(annotate_figure, annotate_figureQ, align = "v", nrow = 2, rel_heights =c(3/7, 4/7), labels = "AUTO")
ggsave(final_figure, file = "forestplot_final.pdf", width = 10, height = 9)

ggsave(filename = "forestplot_final.eps", width = 10, height = 9,
       plot = print(final_figure),
       device = "eps")
```




