---
title: "Effectiveness of Ferritin‐guided Donation Intervals in Blood Donors: results of the Stepped Wedge Cluster‐randomized FIND’EM Trial"
format:
  html:
    code-fold: true
#    page-layout: full
    toc: true
    embed-resources: true
---

```{r, collapse=T}
#| label: load-packages

library(tidyverse)       
library(haven)           # importing SPSS data files
library(furniture)       # nice tables of descriptives
library(corrplot)        # visualize correlations
library(ggplot2)         # plot making
library(GGally)          # extensions to ggplot2
library(sjlabelled)      # work with SPSS data
library(lubridate)       # dates
library(lme4)            # mixed models
library(table1)          # for producing t1
library(sjPlot)          # for model outputs
library(car)             # to calculate VIF
library(broom)           # for function augment

setwd("~/Amber/FINDEM") 
```

# Data

We load and clean the data and present a summary.

```{r, collapse = T}
#| label: process data
data_raw <- readRDS("~/Amber/Data/FINDEM/FINDEM_data.rds")
#clean data
data_clean <- data_raw%>% 
  mutate(HbDef = factor(case_when(HbGoedgekeurd == 0 ~ 1, HbGoedgekeurd == 1 ~ 0))) %>% 
  mutate(Gender = factor(Geslacht) %>% fct_recode("Male" = "Man", "Female" = "Vrouw")) %>% 
  mutate(Timepoint = case_when(Meetweek == 0 ~ as.Date('2017-11-01'), Meetweek == 1 ~ as.Date('2019-03-01'), Meetweek == 2 ~ as.Date('2019-11-01'), TRUE ~ as.Date('2020-11-23'))) %>% 
  mutate(Step = case_when(stap2 == 1.0 ~ as.Date('2017-11-01'), stap2 == 2.1 ~ as.Date('2018-05-01'), stap2 == 2.2~ as.Date('2018-10-01'), stap2 == 3.1 ~ as.Date('2019-03-01'), stap2 == 3.2 ~ as.Date('2019-10-01'), stap2 == 4.0 ~ as.Date('2019-11-01'))) %>% 
  mutate(Intervention_months = interval(Step, Timepoint) %/% months(1)) %>% 
  mutate(Intervention_months =  if_else(Intervention_months < 0, 0, Intervention_months)) %>% 
  mutate(Intervention = factor(case_when(Intervention_months == 0 ~ 0, TRUE & !is.na(Intervention_months)~ 1)))%>% 
  mutate(PostMeno = factor(case_when(Gender == "Female" & LeeftijdBijDOnatie < 45 ~ 0, Gender== "Female" & LeeftijdBijDOnatie >= 45 ~ 1))) %>%
  rename(Ferritin = Ferritine_FINDEM, Age = LeeftijdBijDOnatie, Weight = Gewicht, Height = Lengte)  %>% 
  mutate(IronDef = factor(case_when(Ferritin >= 15 ~ 0, Ferritin < 15 ~ 1)))%>% 
  dplyr::select(-HbGoedgekeurd, -Geslacht, -Einnummer)%>%
  mutate(Intervention_months2 =  if_else(Intervention_months < 6, 0, Intervention_months))%>%
  mutate(Intervention_months_factor = factor(case_when(Intervention_months == 0 ~ 0, Intervention_months == 1 ~ 0, Intervention_months == 5 ~ 0, Intervention_months == 8 ~ 1, Intervention_months == 10 ~ 1, Intervention_months == 12 ~ 2, Intervention_months == 13 ~ 2,Intervention_months == 16 ~ 2,Intervention_months == 18 ~ 3, Intervention_months == 20 ~ 3, Intervention_months == 24 ~ 4, Intervention_months == 25 ~ 4, Intervention_months == 30 ~ 5, Intervention_months == 36 ~ 6))) %>%
  mutate(COVID = factor(ifelse(Donatiedatum >= "2019-10-01" & Donatiedatum <= "2020-02-01", 1,0))) %>% mutate(Hb = Hb/0.6206)


## Quick view
summary(data_clean)
```


```{r}
#| label: recode data
data_clean<-data_clean%>%mutate(Hb = replace(Hb, Hb == (999/0.6206), NA))

#save the data as an Rdta file
save(data_clean, file = "~/Amber/Data/FINDEM/findem_clean.Rdta")
```

Now, we can have a quick look at the cleaned data:

```{r, warning=FALSE, message=FALSE}
#| label: quick data exploration
#| fig-subcap:
#|   - "Plot by pairs"
#|   - "Correlation plot"
#| layout-ncol: 2
data_clean %>%
  na.omit() %>% 
  dplyr::select(-KeyID, -stap2, -Meetweek, -Cluster, -HbDef, -Timepoint, -Step, -Weight, -Height, -Intervention_months,-COVID, -Donatiedatum, -ReturnOpTijd, -Intervention_months2, -Intervention_months_factor, -Step, -Donorcentrum_Cluster, -Donatiecentrumcode) %>%
  GGally::ggpairs(mapping = aes(fill    = Intervention,
                              col       = Intervention,
                              alpha     = 0.05),
                upper = list(continuous = "smooth",
                             combo      = "facethist",
                             discrete   = "ratio"),
                lower = list(continuous = "cor",
                             combo      = "box",
                             discrete   = "facetbar"),
                title = "Quickly Exploring Data") 

data_clean %>% 
  na.omit() %>% 
  dplyr::select(Hb, Ferritin, Weight, Height, Age, Gender, Intervention) %>%
  mutate(Gender = as.numeric(Gender), Intervention = as.numeric(Intervention)) %>%
  rename(Int = Intervention) %>% 
  cor() %>% 
  corrplot.mixed(lower  = "ellipse",
                           upper  = "number",
                           tl.col = "black")
```

We produce an additional scaled and centered dataset. In this dataset, we additionally present ferritin on a Log10 scale.

```{r}
#| label: scaled dataset
data_scale <- data_clean
data_scale$Age <- scale(data_scale$Age, center=TRUE, scale = FALSE)
data_scale$Weight <- scale(data_scale$Weight, center=TRUE, scale = FALSE)
data_scale$Height <- scale(data_scale$Height, center=TRUE, scale = FALSE)
data_scale$FerritinLog10 <- log10(data_scale$Ferritin) 
```

# Descriptives

We present the data stratified by sex and menopausal status.

```{r}
#| label: preprocess t1 data
#for table 1, we only use participants that were assigned to a cluster. Since we present the baseline characteristics of our study, we select only measurement week 0. 
#we load additional data that counts the number of donations
data_teller <- read_sav("~/Amber/Data/FINDEM/baseline_teller.sav")
data_t1 <- data_clean%>% filter(!is.na(Cluster), Meetweek==0) 
data_t1 <- merge(data_t1, data_teller,  by = "KeyID")
data_t1 <- data_t1 %>% mutate(stap3 = as.factor(stap2), stap3 = recode_factor(stap3, "2.1"="2","2.2"="2", "3.1"="3", "3.2"="3"), stap3 = factor(stap3,levels = c("1", "2", "3", "4")), Ndonation = TotaalOpBaseline, Ndonation = replace(Ndonation, Ndonation>0, 1), Ndonation = recode_factor(Ndonation, `1`="0", `0`= "1"))
```

## Premenopausal females

```{r}
#| label: t1 pre-menopausal females
prefemalest1<-table1::table1(~ Age + Hb + Ferritin + Height + Weight + TotaalOpBaseline + Previous2yr + factor(IronDef) + factor(HbDef) + factor(ReturnOpTijd) + factor(Ndonation) | factor(stap3), data = subset(data_t1, data_t1$Gender=="Female" & data_t1$PostMeno==0), caption = "Premenopausal females", render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"), render.categorical=c(.="FREQ (PCT)"))
prefemalest1
```

## Postmenopausal females

```{r}
#| label: t1 post-menopausal females
postfemalest1<-table1(~ Age + Hb + Ferritin + Height + Weight +TotaalOpBaseline + Previous2yr + factor(IronDef) + factor(HbDef) + factor(Return) + factor(ReturnOpTijd) + factor(Ndonation) | factor(stap3), data = subset(data_t1, data_t1$Gender=="Female" & data_t1$PostMeno==1), caption = "Females", render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"), render.categorical=c(.="FREQ (PCT)"))
postfemalest1
```

## Males

```{r}
#| label: t1 males
malest1<- table1(~ Age + Hb + Ferritin + Height + Weight +TotaalOpBaseline + Previous2yr + factor(IronDef) + factor(HbDef) + factor(Return) + factor(ReturnOpTijd) + factor(Ndonation) | factor(stap3), data = subset(data_t1, data_t1$Gender=="Male"), caption = "Males", render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"), render.categorical=c(.="FREQ (PCT)"))
malest1
```

# Analysis

As defined in the protocol, we will fit linear mixed models with random effects for the donation center clusters. We start with analysis of the primary outcomes (Hb levels, ferritin levels, Hb deferral and iron deficiency). We stratify analyses by sex and for females by menopausal status.

## Linear Mixed Models (unscaled)

### Hb levels

```{r, warnings = FALSE, messages = FALSE}
#| label: LMM hb 
hb_fitM <- lmer(Hb ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_clean, subset = data_clean$Gender == "Male")

hb_fit_preF <- lmer(Hb ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_clean, subset = data_clean$Gender == "Female" & data_clean$PostMeno == 0)

hb_fit_postF <- lmer(Hb ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_clean, subset = data_clean$Gender == "Female" & data_clean$PostMeno == 1)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(hb_fit_preF, hb_fit_postF, hb_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Hb levels", digits.re = 4)
```

### Ferritin levels

```{r, warnings = FALSE, messages = FALSE}
#| label: LMM ferritin 
ferritin_fitM <- lmer(Ferritin ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_clean, subset = data_clean$Gender == "Male")

ferritin_fit_preF <- lmer(Ferritin ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_clean, subset = data_clean$Gender == "Female" & data_clean$PostMeno == 0)

ferritin_fit_postF <- lmer(Ferritin ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_clean, subset = data_clean$Gender == "Female" & data_clean$PostMeno == 1)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(ferritin_fit_preF, ferritin_fit_postF, ferritin_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Ferritin levels", digits.re = 4)
```

We see that there is an issue regarding singularity in the model fit for postmenopausal females.

### Hb deferral

```{r, warnings = FALSE, messages = FALSE}
#| label: LMM hb deferral 

hbdef_fitM <- glmer(formula = HbDef ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_clean, subset = data_clean$Gender == "Female" & data_clean$PostMeno == 0, family = binomial)

hbdef_fit_preF <- glmer(formula = HbDef ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_clean, subset = data_clean$Gender == "Female" & data_clean$PostMeno == 0, family = binomial)

hbdef_fit_postF <- glmer(formula = HbDef ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_clean, subset = data_clean$Gender == "Female" & data_clean$PostMeno == 1, family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(hbdef_fit_preF, hbdef_fit_postF, hbdef_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Hb deferral", digits.re = 4)
```

### Iron deficiency

```{r, warnings = FALSE, messages = FALSE}
#| label: LMM iron deficiency 

ID_fitM <- glmer(formula = IronDef ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_clean, subset = data_clean$Gender == "Female" & data_clean$PostMeno == 0, family = binomial)

ID_fit_preF <- glmer(formula = IronDef ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_clean, subset = data_clean$Gender == "Female" & data_clean$PostMeno == 0, family = binomial)

ID_fit_postF <- glmer(formula = IronDef ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_clean, subset = data_clean$Gender == "Female" & data_clean$PostMeno == 1, family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(hbdef_fit_preF, hbdef_fit_postF, hbdef_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Iron deficiency", digits.re = 4)
```

::: {.callout-note appearance="default"}
### Decision

The basic models for Hb deferral and iron deficiency in males and premenopausal females failed to converge. Because of this, we decided to scale the variables for all analyses.
:::

## Linear Mixed Models (scaled)

### Hb levels

```{r, warnings = FALSE, messages = FALSE}
#| label: LMM hb scaled
hb_fitM <- lmer(Hb ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_scale, subset = data_scale$Gender == "Male")

hb_fit_preF <- lmer(Hb ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0)

hb_fit_postF <- lmer(Hb ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(hb_fit_preF, hb_fit_postF, hb_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Hb levels", digits.re = 4)
```

### Log10 Ferritin levels

```{r, warnings = FALSE, messages = FALSE}
#| label: LMM ferritin scaled
ferritin_fitM <- lmer(FerritinLog10 ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_scale, subset = data_scale$Gender == "Male")

ferritin_fit_preF <- lmer(FerritinLog10 ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0)

ferritin_fit_postF <- lmer(FerritinLog10 ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(ferritin_fit_preF, ferritin_fit_postF, ferritin_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Log10 Ferritin levels", digits.re = 4)
```

### Hb deferral

```{r, warnings = FALSE, messages = FALSE}
#| label: LMM hb deferral scaled

hbdef_fitM <- glmer(formula = HbDef ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0, family = binomial)

hbdef_fit_preF <- glmer(formula = HbDef ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0, family = binomial)

hbdef_fit_postF <- glmer(formula = HbDef ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1, family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(hbdef_fit_preF, hbdef_fit_postF, hbdef_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Hb deferral", digits.re = 4)
```

### Iron deficiency

```{r, warnings = FALSE, messages = FALSE}
#| label: LMM iron deficiency scaled

ID_fitM <- glmer(formula = IronDef ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0, family = binomial)

ID_fit_preF <- glmer(formula = IronDef ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0, family = binomial)

ID_fit_postF <- glmer(formula = IronDef ~ Age + Weight + Height + Intervention_months_factor + (1|Cluster), data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1, family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(hbdef_fit_preF, hbdef_fit_postF, hbdef_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Iron deficiency", digits.re = 4)
```

### Model summaries

Below we present the complete model summaries. Here, the sections "Random effects" are of interest, as these statistics were not presented in the above tables.

```{r, collapse=T}
#| label: model summaries
summary(hb_fitM)
summary(hb_fit_preF)
summary(hb_fit_postF)

summary(ferritin_fitM)
summary(ferritin_fit_preF)
summary(ferritin_fit_postF)
```

::: {.callout-note appearance="default"}
### Decision

We decide to proceed with the analyses using standard linear regression models using only fixed effects, thus disregarding the random effects from clustering by donation center cluster. We decided to deviate from the protocol based on:

-   the ICC generally low ICCs that range from 0.000 to 0.01

-   the very small difference between the marginal and conditional R squared

-   the variance between the clusters is smaller than the residual variance

Which indicates that the random effects of the donation center clusters in these analyses are of minimal importance and can be disregarded.
:::

## Linear Regression
### Hb levels
```{r}
#| label: lin reg Hb levels
hb_fitM <- lm(Hb ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male")
hb_fit_preF <- lm(Hb ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0)
hb_fit_postF <- lm(Hb ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(hb_fit_preF, hb_fit_postF, hb_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Hb levels")
```

### Ferritin levels
```{r}
#| label: lin reg ferritin levels
ferritin_fitM <- lm(FerritinLog10 ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male")
ferritin_fit_preF <- lm(FerritinLog10 ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0)
ferritin_fit_postF <- lm(FerritinLog10 ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(ferritin_fit_preF, ferritin_fit_postF, ferritin_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Ferritin levels")
```

## Logistic Regression
### Hb deferral

```{r, message=F, error=F}
#| label: log reg Hb deferral

hbdef_fitM <- glm(formula = HbDef ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male" ,family = binomial)

hbdef_fit_preF <- glm(formula = HbDef ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0,family = binomial)

hbdef_fit_postF <- glm(formula = HbDef ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1,family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(hbdef_fit_preF, hbdef_fit_postF, hbdef_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Hb deferral")
```
### Iron deficiency
```{r, message=F, error=F}
#| label: log reg iron deficiency

ID_fitM <- glm(formula = IronDef ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male" ,family = binomial)

ID_fit_preF <- glm(formula = IronDef ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0,family = binomial)

ID_fit_postF <- glm(formula = IronDef ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1,family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(ID_fit_preF, ID_fit_postF, ID_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Iron deficiency")
```

### Donor return (on time)
```{r, message=F, error=F}
#| label: log reg return on time
returnOT_fitM <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male",family = binomial)
returnOT_fit_preF <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0,family = binomial)
returnOT_fit_postF <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1,family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(returnOT_fit_preF, returnOT_fit_postF, returnOT_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Donor return (on time)")
```

### Donor return adjusted for COVID
```{r, message=F, error=F}
#| label: log reg return on time including covid
returnOT_fitM <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + COVID + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male",family = binomial)
returnOT_fit_preF <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + COVID + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0,family = binomial)
returnOT_fit_postF <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + COVID + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1,family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(returnOT_fit_preF, returnOT_fit_postF, returnOT_fitM, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Donor return (on time)")
```

## Model assumptions

### Hb levels
```{r}
#| label: assumptions Hb 
par(mfrow = c(2, 2))
plot(hb_fit_preF)
plot(hb_fit_postF)
plot(hb_fitM)
```

### Ferritin levels
```{r}
#| label: assumptions log10 ferritin
par(mfrow = c(2, 2))
plot(ferritin_fit_preF)
plot(ferritin_fit_postF)
plot(ferritin_fitM)
```

### Hb deferral
Independent observations: yes, we checked clustering with the LMM. Moreover, there is no longitudinal dependence. 

Influential observations:
```{r}
#| label: hb deferral assumptions

plot(hbdef_fit_preF, which = 4) #this one has one quite curious observation so i want to look at the standardized residual
model.data <- augment(hbdef_fit_preF) %>% 
  dplyr::mutate(index = 1:n()) 
model.data %>% top_n(3, .cooksd) #the absolute standardized residual is not above 3 so we don't have to worry about this outlier.
```

Since this observation has a quite curious cook's distance, we looked at the absolute standardized residual. However, it's not above 3 so we don't have to worry about it. 

```{r}

plot(hbdef_fit_postF, which = 4)
plot(hbdef_fitM, which = 4)

#no multicollinearity: 

vif(hbdef_fit_preF)
vif(hbdef_fit_postF)
vif(hbdef_fitM)
```

VIF is not above 4 for any of the covariates


### Iron deficiency
```{r}
#| label: iron deficiency assumptions
#independent observations: yes, we checked clustering with the LMM and there is no longitudinal dependence. 

#influential observations:
plot(ID_fit_preF, which = 4) 
plot(ID_fit_postF, which = 4)
plot(ID_fitM, which = 4)

#no multicollinearity: 

vif(ID_fit_preF)
vif(ID_fit_postF)
vif(ID_fitM)
```
VIF is not above 4 for any of the covariates

### Donor return
```{r}
#| label: donor return assumptions
#independent observations: yes, we checked clustering with the LMM and there is no longitudinal dependence. 

#influential observations:
plot(returnOT_fit_preF, which = 4) 
model.data <- augment(returnOT_fit_preF) %>% 
  dplyr::mutate(index = 1:n()) 
model.data %>% top_n(3, .cooksd) 

plot(returnOT_fit_postF, which = 4)

model.data <- augment(returnOT_fit_postF) %>% 
  dplyr::mutate(index = 1:n()) 
model.data %>% top_n(3, .cooksd) 

```

Again some observations that stand out. However, the absolute standardized residuals are not above 3.

```{r}

plot(returnOT_fitM, which = 4)

#no multicollinearity: 

vif(returnOT_fit_preF)
vif(returnOT_fit_postF)
vif(returnOT_fitM)
```

VIF is not above 4 for any of the covariates

::: {.callout-note appearance="default"}
### Decision

We conclude that the assumptions for none of the models were violated.
:::
