---
title: "Effectiveness of Ferritin‐guided Donation Intervals in Blood Donors: results of the Stepped Wedge Cluster‐randomized FIND’EM Trial"
format:
  html:
    code-fold: true
#    page-layout: full
    toc: true
    embed-resources: true
---

```{r}
#| label: load-packages
#| include: false

library(tidyverse)       
library(haven)           # importing SPSS data files
library(furniture)       # nice tables of descriptives
library(corrplot)        # visualize correlations
library(ggplot2)         # plot making
library(GGally)          # extensions to ggplot2
library(sjlabelled)      # work with SPSS data
library(lubridate)       # dates
library(lme4)            # mixed models
library(table1)          # for producing t1
library(sjPlot)          # for model outputs
library(car)             # to calculate VIF
library(broom)           # for function augment
library(stringr)

setwd("~/Amber/FINDEM") 
```

# Data

We load and clean the data and present a summary.

```{r, collapse = T}
#| label: process data
data_raw <- readRDS("data/FINDEM_data.rds")
#clean data
data_clean <- data_raw%>% 
  mutate(HbDef = factor(case_when(HbGoedgekeurd == 0 ~ 1, HbGoedgekeurd == 1 ~ 0))) %>% 
  mutate(Gender = factor(Geslacht) %>% fct_recode("Male" = "Man", "Female" = "Vrouw")) %>% 
  mutate(Timepoint = case_when(Meetweek == 0 ~ as.Date('2017-11-01'), Meetweek == 1 ~ as.Date('2019-03-01'), Meetweek == 2 ~ as.Date('2019-11-01'), TRUE ~ as.Date('2020-11-23'))) %>% 
  mutate(Step = case_when(stap2 == 1.0 ~ as.Date('2017-11-01'), stap2 == 2.1 ~ as.Date('2018-05-01'), stap2 == 2.2~ as.Date('2018-10-01'), stap2 == 3.1 ~ as.Date('2019-03-01'), stap2 == 3.2 ~ as.Date('2019-10-01'), stap2 == 4.0 ~ as.Date('2019-11-01'))) %>% 
  mutate(Intervention_months = interval(Step, Timepoint) %/% months(1)) %>% 
  mutate(Intervention_months =  if_else(Intervention_months < 0, 0, Intervention_months)) %>% 
  mutate(Intervention = factor(case_when(Intervention_months == 0 ~ 0, TRUE & !is.na(Intervention_months)~ 1)))%>% 
  mutate(PostMeno = factor(case_when(Gender == "Female" & LeeftijdBijDOnatie < 45 ~ 0, Gender== "Female" & LeeftijdBijDOnatie >= 45 ~ 1))) %>%
  rename(Ferritin = Ferritine_FINDEM, Age = LeeftijdBijDOnatie, Weight = Gewicht, Height = Lengte)  %>% 
  mutate(IronDef = factor(case_when(Ferritin >= 15 ~ 0, Ferritin < 15 ~ 1)))%>% 
  dplyr::select(-HbGoedgekeurd, -Geslacht)%>%
  mutate(Intervention_months2 =  if_else(Intervention_months < 6, 0, Intervention_months))%>%
  mutate(Intervention_months_factor = factor(case_when(Intervention_months == 0 ~ 0, Intervention_months == 1 ~ 0, Intervention_months == 5 ~ 0, Intervention_months == 8 ~ 1, Intervention_months == 10 ~ 1, Intervention_months == 12 ~ 2, Intervention_months == 13 ~ 2,Intervention_months == 16 ~ 2,Intervention_months == 18 ~ 3, Intervention_months == 20 ~ 3, Intervention_months == 24 ~ 4, Intervention_months == 25 ~ 4, Intervention_months == 30 ~ 5, Intervention_months == 36 ~ 6))) %>%
  mutate(COVID = factor(case_when(Donatiedatum < "2019-10-01" ~ 0, Donatiedatum >= "2019-10-01" ~ 1)))

summary(data_clean)

data_clean<-data_clean%>%mutate(Hb = replace(Hb, Hb == 999, NA))%>%filter(!is.na(Cluster))

```

# Table 1

See section Descriptives of main FINDEM.html file.

# Table 2

See section Analysis \> Linear regression of main FINDEM.html file.

# Table 3

See section Analysis of FINDEM_questionnaire.html file.

# Table S1

See section Descriptives of FINDEM_questionnaire.html file.

# Table S2

For this sensitivity analysis, we need to scale the data:

```{r}
#| label: scaled dataset
data_scale <- data_clean
data_scale$Age <- scale(data_scale$Age, center=TRUE, scale = FALSE)
data_scale$Weight <- scale(data_scale$Weight, center=TRUE, scale = FALSE)
data_scale$Height <- scale(data_scale$Height, center=TRUE, scale = FALSE)
data_scale$FerritinLog10 <- log10(data_scale$Ferritin) 
```

For ferritin $\le$ 30:

```{r}
return_fitM_b30 <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male" & data_scale$Ferritin <= 30,family = binomial)
return_fit_preF_b30 <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0 & data_scale$Ferritin <= 30,family = binomial)
return_fit_postF_b30 <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1 & data_scale$Ferritin <= 30,family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(return_fit_preF_b30, return_fit_postF_b30, return_fitM_b30, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Donor return (on time)")
```

For ferritin \> 30:
```{r}
return_fitM_a30 <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male" & data_scale$Ferritin > 30,family = binomial)
return_fit_preF_a30 <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0 & data_scale$Ferritin > 30,family = binomial)
return_fit_postF_a30 <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1 & data_scale$Ferritin > 30,family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(return_fit_preF_a30, return_fit_postF_a30, return_fitM_a30, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Donor return (on time)")
```

# Table S3

To give a clear overview of the differences between the pre- and post-implementation situation in the donation centers, we present the following table:

```{r}
SupTable3<-table1::table1(~ Age + Hb + Ferritin + Height + Weight + factor(IronDef) + factor(HbDef) + factor(ReturnOpTijd) | factor(Meetweek), data = subset(data_clean, data_clean$Meetweek =="1" | data_clean$Meetweek=="4"), caption = "Premenopausal females", render.continuous=c(.="Mean (SD)", .="Median [Q1, Q3]"), render.categorical=c(.="FREQ (PCT)"), overall=F)
SupTable3
```

# Table S4

We have conducted a sensitivity analysis on the donor return outcome, separating donors that were deferred based on Hb and donors that were deferred based on ferritin levels.

For donors that were deferred based on Hb levels: 

```{r}
return_fitM_HbDef <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male" & data_scale$HbDef == 1,family = binomial)
return_fit_preF_HbDef <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0 & data_scale$HbDef == 1,family = binomial)
return_fit_postF_HbDef <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1 & data_scale$HbDef == 1,family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(return_fit_preF_HbDef, return_fit_postF_HbDef, return_fitM_HbDef, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Donor return (on time) for Hb deferred donors")
```

For donors that were deferred based on ferritin levels:

```{r}
return_fitM_b30 <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Male" & data_scale$Ferritin <= 30,family = binomial)
return_fit_preF_b30 <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 0 & data_scale$Ferritin <= 30,family = binomial)
return_fit_postF_b30 <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale, subset = data_scale$Gender == "Female" & data_scale$PostMeno == 1 & data_scale$Ferritin <= 30,family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(return_fit_preF_b30, return_fit_postF_b30, return_fitM_b30, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Donor return (on time) for donors whose donation intervals were extended")
```

# Table S5

We stratify the analysis of the donor return outcome by urban vs. rural donation locations. Urban locations are defined as fixed donation centers, rural locations are defined as mobile donation centers. 

```{r}
#we first need to find whether a location was done at a mobile location. We find this through linking our data to the donation database.
alldonations <- readRDS("~/Amber/Data/alledonaties_2008_2022.rds")
alldonations <- subset(alldonations, select = c("EINnummer", "Donatiecentrumcode"))
alldonations$EINnummer <- substr(alldonations$EINnummer,1,13)
data_scale2 <- merge(data_scale, alldonations, by.x = "Einnummer", by.y = "EINnummer", all.x = T)
data_scale2 <- data_scale2 %>% mutate(MAL = grepl("ML", Donatiecentrumcode)) %>% mutate(MAL = case_when(MAL == FALSE ~ 0,MAL == TRUE ~ 1 ))
```

For donors that located in an urban center:
```{r}
return_fitM_AL <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale2, subset = data_scale2$Gender == "Male" & data_scale2$MAL ==0,family = binomial)
return_fit_preF_AL <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale2, subset = data_scale2$Gender == "Female" & data_scale2$PostMeno == 0 & data_scale2$MAL==0,family = binomial)
return_fit_postF_AL <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale2, subset = data_scale2$Gender == "Female" & data_scale2$PostMeno == 1 & data_scale2$MAL==0,family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(return_fit_preF_AL, return_fit_postF_AL, return_fitM_AL, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Donor return (on time) for donors at an urban locatoin")
```

For donors that donated at a rural location:

```{r}
return_fitM_MAL <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale2, subset = data_scale2$Gender == "Male" & data_scale2$MAL ==1,family = binomial)
return_fit_preF_MAL <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale2, subset = data_scale2$Gender == "Female" & data_scale2$PostMeno == 0 & data_scale2$MAL==1,family = binomial)
return_fit_postF_MAL <- glm(formula = ReturnOpTijd ~ Age + Weight + Height + Intervention_months_factor, data = data_scale2, subset = data_scale2$Gender == "Female" & data_scale2$PostMeno == 1 & data_scale2$MAL==1,family = binomial)

pl <- c(
  `(Intercept)` = "Intercept",
  Intervention_months_factor1 = "6-11 months",
  Intervention_months_factor2 = "12-17 months",
  Intervention_months_factor3 = "18-23 months",
  Intervention_months_factor4 = "24-29 months",
  Intervention_months_factor5 = "30-35 months",
  Intervention_months_factor6 = "36+ months"
)

tab_model(return_fit_preF_MAL, return_fit_postF_MAL, return_fitM_MAL, pred.labels = pl, dv.labels = c("Pre-menopausal women", "Post-menopausal women", "Men"), show.reflvl = T, title = "Donor return (on time) for donors at a rural location")
```
# Figure 1

Figure 1 was produced in Microsoft Powerpoint.

# Figure 2

```{r}
data_clean$LowFerr <- 0
data_clean$LowFerr[data_clean$Ferritin>=15 & data_clean$Ferritin < 30] <- 1

figure2M <- data.frame(matrix(NA,    # Create empty data frame
                          nrow = 7, # we make a dataset per sex and then have one row per month group of the intervention
                          ncol = 4)) #we need columns for the percentage hb deferral, iron deficiency and low ferritin
colnames(figure2M) <- c("Months", "Hb", "ID", "LowFer")
figure2M$Months <- c("0-5 months", "6-11 months", "12-17 months", "18-23 months", "24-29 months", "30-35 months", "36+ months")

figure2PreF <- data.frame(matrix(NA,    # Create empty data frame
                          nrow = 7, # we make a dataset per sex and then have one row per month group of the intervention
                          ncol = 4)) #we need columns for the percentage hb deferral, iron deficiency and low ferritin
colnames(figure2PreF) <- c("Months", "Hb", "ID", "LowFer")
figure2PreF$Months <- c("0-5 months", "6-11 months", "12-17 months", "18-23 months", "24-29 months", "30-35 months", "36+ months")

figure2PostF <- data.frame(matrix(NA,    # Create empty data frame
                          nrow = 7, # we make a dataset per sex and then have one row per month group of the intervention
                          ncol = 4)) #we need columns for the percentage hb deferral, iron deficiency and low ferritin
colnames(figure2PostF) <- c("Months", "Hb", "ID", "LowFer")
figure2PostF$Months <- c("0-5 months", "6-11 months", "12-17 months", "18-23 months", "24-29 months", "30-35 months", "36+ months")

for (i in 1:7){
  figure2M[i,2] <- (length(which(data_clean$Gender == "Male" & data_clean$Intervention_months_factor == (i-1) & data_clean$HbDef==1))/length(which(data_clean$Gender == "Male" & data_clean$Intervention_months_factor == (i-1))))*100
  figure2M[i,3] <- (length(which(data_clean$Gender == "Male" & data_clean$Intervention_months_factor == (i-1) & data_clean$IronDef==1))/length(which(data_clean$Gender == "Male" & data_clean$Intervention_months_factor == (i-1))))*100
    figure2M[i,4] <- (length(which(data_clean$Gender == "Male" & data_clean$Intervention_months_factor == (i-1) & data_clean$LowFerr ==1))/length(which(data_clean$Gender == "Male" & data_clean$Intervention_months_factor == (i-1))))*100
    
      figure2PreF[i,2] <- (length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 0 & data_clean$Intervention_months_factor == (i-1) & data_clean$HbDef==1))/length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 0 & data_clean$Intervention_months_factor == (i-1))))*100
  figure2PreF[i,3] <- (length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 0 & data_clean$Intervention_months_factor == (i-1) & data_clean$IronDef==1))/length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 0 & data_clean$Intervention_months_factor == (i-1))))*100
    figure2PreF[i,4] <- (length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 0 & data_clean$Intervention_months_factor == (i-1) & data_clean$LowFerr ==1))/length(which(data_clean$Gender == "Female" &  data_clean$PostMeno == 0 & data_clean$Intervention_months_factor == (i-1))))*100
    
          figure2PostF[i,2] <- (length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 1 & data_clean$Intervention_months_factor == (i-1) & data_clean$HbDef==1))/length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 1 & data_clean$Intervention_months_factor == (i-1))))*100
  figure2PostF[i,3] <- (length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 1 & data_clean$Intervention_months_factor == (i-1) & data_clean$IronDef==1))/length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 1 & data_clean$Intervention_months_factor == (i-1))))*100
    figure2PostF[i,4] <- (length(which(data_clean$Gender == "Female" & data_clean$PostMeno == 1 & data_clean$Intervention_months_factor == (i-1) & data_clean$LowFerr ==1))/length(which(data_clean$Gender == "Female" &  data_clean$PostMeno == 1 & data_clean$Intervention_months_factor == (i-1))))*100
}


```